---
title: "Final Project"
output: html_notebook
Name: 'Daniel Lee'
---

### IDEA & GOAL

The main goal of this final project is to analyze how variables such as age, batting average and salary of the players are related by joining two different tables which are 'Player Standard Batting' and 'Salary'. These two tables will be analyzed separately and joined together to look for the relationship between the variables by plotting graphs. 

### Setup
```{r}
rm(list = ls())
library(lubridate)
library(tidyverse)
library(mosaic)
library(rvest)
```

### Loading datasets
```{r}
# Loading two tables: 'Player Standard Batting' & 'Salary'
library(readxl)
Player_Standard_Batting <- read_excel("~/Desktop/Penn State/Courses/STAT 184/Final Project/Player Standard Batting.xlsx")
View(Player_Standard_Batting)

library(readxl)
Salary <- read_excel("~/Desktop/Penn State/Courses/STAT 184/Final Project/Salary.xlsx")
View(Salary)
```

```{r}
# Removing unnecessary punctuations at the name column for 'Salary' table
Salary <-
  Salary %>%
  mutate(Name = gsub(pattern = "[[:punct:]]", replacement = "", x = Name))

glimpse(Salary)
```


### Data Wrangling for 'Player Standard Batting' table
```{r}
# Creating a new table (PSB) with variables that are necessary and significant & Removing unnecessary punctuations at the name column for 'Player Standard Batting' table
PSB <- 
  Player_Standard_Batting %>%
  filter(POS != "1") %>% 
  select(Name, Tm, Lg, Age, R, H, HR, RBI, AB, BA, POS) %>%
  mutate(Name = gsub(pattern = "[[:punct:]]", replacement = "", x = Name))

View(PSB)

# Average Hits and Batting average
PSB %>%
  summarise(BA = mean(BA, na.rm = TRUE) , H = mean(H, na.rm = TRUE)) # Pitchers Batting Average excluded 
```

```{r}
# Most Hits and Batting aveage for players who were at bat for more or equal to 300 times. 
PSB %>%
  filter(AB >= 300) %>%        # Filtering out players who were not at bat less than 300 games
  summarise(BA = max(BA, na.rm = TRUE), H = max(H, na.rm = TRUE))
```


```{r}
# Counting the number of players who are 27 years old in each team & Removing pitchers and duplicated names from the data table occured due to trade during season.
PSB %>%
  group_by(Tm) %>%
  filter(Age >= 27) %>%
  filter(POS != "1") %>%                     # Pitchers have a number position = 1 
  summarise(count = n( )) %>%
  arrange(desc(count))
```

```{r}
# Plotting a graph: Age vs BA into two groups (AL & NL)
PSB %>%
  ggplot(aes(x = Age, y = BA, color = Lg)) +
  stat_smooth() +
  ylab("Batting Average (BA)") + xlab("Age (years)")
```
The graph (figure 1.1) shows that, in average, batters in American league have higher batting average betwen the age of 21 to 37 years old which are most of the players. However, this graph can be also interpreted as national league batters who are younger than 21 years old having more potential. Batting average for national league players could be higher than american league players in few years. The following graph could be altered next year because there are many trades happening between the teams from National league and American league. 

```{r}
# Plotting a graph: Age vs H into two groups (AL & NL)
PSB %>%
  ggplot(aes(x = Lg, y = H)) +
  geom_boxplot()+
  ylab("Hits (H)") + xlab("League (Lg)")
```
This boxplot (figure 1.2) shows how number of hits affect the above graph (BA vs Age). The American League has higher Q1, Q3 and median than National league's Q1, Q3 and median. Although hits are not 100% the factor of having a higher batting average as shown in the figure 1.1, but it is a significant factor.

### Data Wrangling for 'Salary' table
```{r}
# Finding the average salary in 2018 
Salary %>%
  summarise(AVG = mean(AVG, na.rm = TRUE))

# Number of players who earned more than the average salary ($4,520,000) for each team in 2018
Salary %>%
  group_by(Tm) %>%
  filter(AVG > 4520000) %>%
  filter(POS != "1") %>%         # Pitchers have a number position = 1 / POS does not work 
  summarise(count = n( )) %>%
  arrange(desc(count))
```

```{r}
# Highest and lowest salary for each position
Salary %>%
  group_by(POS) %>%
  summarise(AVG = max(AVG, na.rm = TRUE)) +
  arrange(desc())

```


```{r}
ggplot(data = Salary, aes(x = AVG)) + 
  geom_histogram(binwidth = 1400000) + 
  aes(colour = POS) + 
  xlim=c(0, 35000000), ylim=c(0, 300)

Salary %>%
  ggplot(aes(x = AVG)) +
  geom_density(color = "black", fill = "gray", alpha = 0.75) +
  xlab("SALARY") +
  xlim(545000,30714286)

PSB %>%
  ggplot(aes(x = Age, y = HR, color = Lg)) +
  geom_point() +
  ylab("Batting Average (BA)") + xlab("Age (years)")
```


### Joining two tables = 'Player Standard Batting' & 'Salary'
```{r}
PSB <-
  PSB %>%
  mutate(Name = ungroup(Name)) %>%
  mutate(Tm = ungroup(Tm)) 

Salary <-
  Salary %>%
  mutate(Name = ungroup(Name)) %>%
  mutate(Tm = ungroup(Tm)) 


MLB <-
  PSB %>%
  left_join(Salary, by = c("Name", "Tm"))
View(MLB)


```

### Plotting a graph to find the relationship between age vs salary & BA vs salary
```{r}

```

